---
title: "Lab5: Spatial interpolation of your own data"
author: "Katarzyna Goch"
date: "24 Nov 2018"
output: html_document
---

This is the report for the fifth assignment "Spatial interpolation of your own dataset" from course "Analysis of Spatio-Temporal Data", WS 2018/19.  
Supervisor: prof. Edzer Pebesma

```{r setup, include=FALSE}

library(sp)
library(gstat)
library(rgdal)
library(mapview)

```
  
****************


## Import your own geostatistical dataset in R
I am using the sama data as in Assignment 04.

* Kashparov, V.; Levchuk, S.; Zhurba, M.; Protsak, V.; Khomutinin, Yu.; Beresford, N.A.; Chaplow, J.S. (2017). Spatial datasets of radionuclide contamination in the Ukrainian Chernobyl Exclusion Zone. NERC Environmental Information Data Centre. https://doi.org/10.5285/782ec845-2135-4698-8881-b38823e533bf
  
```{r, fig.align='center', fig.width=9}
# Read in the data
df1 <- read.csv2("Chernobyl_pH.csv",
                header=TRUE,
                quote="\"",
                stringsAsFactors = FALSE,
                strip.white = TRUE)

# Get Longitue and latitude data from data frame
myPoints <- df1
myPoints$Latitude<- as.numeric(as.character(df1$Latitude))
myPoints$Longitude<- as.numeric(as.character(df1$Longitude))
myPoints$pH_H20<- as.numeric(as.character(df1$pH_H20))
coordinates(myPoints)<-~Longitude+Latitude
proj4string(myPoints) <- CRS("+init=epsg:4326")
ph<-spTransform(myPoints,CRS=CRS("+init=epsg:32635"))

# Remove overlapping measurement
ph <- ph[-zerodist(ph)[,1],]

mapview(ph,zcol="pH_H20", at=seq(3,9,1), map.types = "OpenStreetMap", legend=TRUE, layer.name=c("pH in H2O"), cex="pH_H20")

# Fit for the data from the 50 km cutoff
v <- variogram(pH_H20~1,data=ph[!is.na(ph$pH_H20),], cutoff = 20000, width=200)
v.fit = fit.variogram(v, vgm(0.6,"Exp",3000))
plot(v, v.fit)

```
  
## Create a prediction grid (see lecture slides) covering your study area
```{r, fig.align='center', fig.width=9}
# Create a grid
x = seq(663591, 737845, 1000)
y = seq(5649073,	5715455, 1000)
grd = expand.grid(x = x, y = y)
head(grd)
coordinates(grd) <-~x+y
gridded(grd)=TRUE
proj4string(grd) <- CRS("+init=epsg:32635") 
summary(grd)

```
  
## Interpolate your variable using the variogram model you fitted in the previous assignment
```{r, fig.align='center'}
# model used:
v.fit

# Krige
k = krige(pH_H20~1, ph[!is.na(ph$pH_H20),], grd, v.fit)
summary(k["var1.pred"])

# Save the results
sum1<-capture.output(summary(k["var1.pred"])[7])
```
  
## Visualise the results
```{r, fig.align='center', fig.width=9}
# Visualize
spplot(k["var1.pred"], main = "pH in H2O predictions. Model: Exp, sill: 0.607, range: 222")

```
  
## Experiment with changing the variogram model: increase and decrease the nugget variance, increase and decrease the range parameter, and discuss how the resulting map changes.
  
**1. Range changed from 222 m to 500 m**
  
```{r, fig.width=9}
# Chcange the range of the model to 10 km
v.fit = vgm(0.607,"Exp",500)
# Krige
k = krige(pH_H20~1, ph[!is.na(ph$pH_H20),], grd, v.fit)
spplot(k["var1.pred"], main = "pH in H2O predictions. Model: Exp, still: 0.607, range: 500 m")

# Save results
sum2<-capture.output(summary(k["var1.pred"])[7])

```

  
**2. Range changed to 1 km**
  
```{r, fig.width=9}
# Chcange the model to Sph
v.fit = vgm(0.607,"Exp",1000)
# Krige
k = krige(pH_H20~1, ph[!is.na(ph$pH_H20),], grd, v.fit)
spplot(k["var1.pred"], main = "pH in H2O predictions. Model: Exp, still: 0.607, range: 1 km")

#Save results
sum3<-capture.output(summary(k["var1.pred"])[7])
```

 
 **3. Range not changed, but sill changed to 0.2**
  
```{r, fig.width=9}
# Chcange the model to Sph
v.fit = vgm(0.2,"Exp",222)
# Krige
k = krige(pH_H20~1, ph[!is.na(ph$pH_H20),], grd, v.fit)
spplot(k["var1.pred"], main = "pH in H2O predictions. Model: Exp, still: 0.2, range: 222 m")

#Save results
sum4<-capture.output(summary(k["var1.pred"])[7])
``` 
**4. Compare the results**
  
```{r}
# Comprae the results
out_comp <- matrix(c(sum1[3],sum2[3],sum3[3], sum4[3],sum1[8],sum2[8],sum3[8], sum4[8], sum1[5],sum2[5],sum3[5], sum4[5]),ncol=4,byrow=TRUE)
colnames(out_comp) <- c("'Base model","Range 500 m","Range 1 km", "Sill 0.2")
as.table(out_comp)

```
We can see the differece in the output values when the range was changed. No difference was observed, when the still was amended (?).
  
  
## (bonus:) try to create a set of (say 10) conditional simulations for your dataset. How do these differ from interpolations?